Analyzing Strawberry Pricing Strategies Across Canadian Vendors\thanks{Code and data are available at: https://github.com/HeTianning/The-price-of-strawberries}

### 2024-12-03 paper/paper.tex
Improve and revise the paper

---
title: "Analyzing Strawberry Pricing Strategies Across Canadian Vendors"
subtitle: "Price Variations Driven by Supplier Strategies and Seasonal Demand in Canadian Grocery Markets"
author: 
  - Tianning He
thanks: "Code and data are available at: https://github.com/HeTianning/The-price-of-strawberries"
date: 25 November 2024
date-format: long
abstract: "This study investigates the pricing of strawberries across eight major suppliers in Canada: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods. Using a dataset containing variables such as current price, old price, and month, a Bayesian regression model was applied to analyze the relationships between pricing strategies, seasonal trends, and vendor-specific practices. Results reveal significant price variation among suppliers, with old price showing a strong positive correlation with current price, reflecting the impact of prior pricing strategies on current prices. Seasonal trends captured by the month variable indicate subtle but consistent fluctuations in pricing, highlighting the influence of supply and demand dynamics. These findings provide actionable insights into supplier-specific pricing strategies and market behavior, offering valuable guidance to consumers and stakeholders in the Canadian strawberry market."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(ggplot2)
library(dplyr)
library(arrow)
library(here)
library(modelsummary)
library(bayesplot)
library(rstanarm)
```

# Introduction

Strawberries are one of the most important small berries in Canada and the United States(@macoun1919strawberry), and pricing of strawberries in the Canadian retail market varies widely due to factors such as supplier strategies, promotional practices, and temporal trends. This investigation examines the determinants of strawberry prices, focusing on differences between different suppliers, the role of discounts, and seasonal effects. By utilizing a Bayesian linear regression model, the study examines the relationship between key predictors (supplier, original price, promotion, and time) and the current price of strawberries. The Bayesian approach not only provides parameter estimates but also quantifies uncertainty, thus providing a nuanced understanding of price dynamics in competitive and seasonal markets.

The main objective of this analysis is to estimate and explain the factors that influence retail strawberry prices in Canada. The dependent variable current price reflects the Canadian dollar price per pound of strawberries. The key predictors include supplier identity (categorical), which captures supplier-specific pricing strategies; old price (continuous), which represents the original price before promotion, and month (continuous), which captures temporal or seasonal trends(@hodgdon2024current). By modeling these relationships, the study aims to quantify how these variables affect pricing and explore their combined impact on strawberry price changes.

@@

The rest of the paper is structured as follows. @sec-data describes the data collection process, including data sources, sampling strategy, and preprocessing steps taken to ensure data quality. @sec-model outlines the methodological framework, focusing on the Bayesian linear regression model, its formulation, and diagnostic checks. @sec-result presents the main findings of the analysis, focusing on specific supplier pricing strategies, seasonal trends, and the impact of sales and promotions on strawberry prices. @sec-discussion discusses the implications of these findings, placing them in the broader context of consumer behavior and supplier strategies in the fresh produce market.

# Data {#sec-data}

To conduct the analysis, `R`, a versatile statistical programming language, was utilized for simulating, testing, downloading, cleaning, and modeling data [@citeR]. Several powerful libraries supported various stages of the analysis. The `tidyverse` package suite [@tidyverse], which includes `dplyr` for data manipulation and `ggplot2` for visualization, was central to handling and exploring the dataset. Data cleaning and structuring were enhanced using `janitor`, while `arrow` facilitated efficient handling of large datasets by enabling seamless reading and writing of data [@arrow]. The `here` package streamlined file management, ensuring robust and reproducible workflows [@here].

Visualization and reporting were further supported by `knitr` [@knitr], which enabled dynamic report generation, and `modelsummary` [@modelsummary], which provided professional-quality summaries for statistical models. Bayesian modeling and advanced statistical analysis were conducted using `rstanarm` [@rstanarm], while posterior diagnostics and visualizations were enhanced with `bayesplot` [@bayesplot]. Testing and quality assurance of code were carried out with `testthat` [@testthat], ensuring the reliability and reproducibility of results.

@@

## Overview

The data for this analysis was provided by Project Hammer [@filipp2024hammer], which was collected through a process of scraping website UI screens from vendors within the Toronto community. It contains data on product details (e.g., product name, brand, current and old price, and unit of product) for 8 different grocery stores (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods). Project Hammer's data is available from February 28, 2024, and since the data is collected by extracting information from grocery store websites, there may be missing data, as discussed in appendix. Through the method of screen scraping, the raw data is intended to measure and reflect grocery store vendor decisions and product pricing trends.

For this analysis, the eight vendors - Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods - and their pricing for strawberries were considered. The variable “price” for each supplier in this analysis reflects the current price of their product at the time the data was downloaded.

## Measurement

@@

The dataset also incorporates vendor as a categorical variable, identifying one of eight major suppliers: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods. This variable is pivotal for inter-vendor comparisons, enabling the analysis to uncover unique pricing strategies, competitive behaviors, and market segmentation. By examining differences in pricing across these suppliers, the analysis can identify which vendors consistently offer lower prices, adopt aggressive discounting practices, or align their pricing with premium branding.

@@

## Outcome variable

@@

By examining both current price and old price, the study provides a detailed view of promotional activity and vendor behavior. A significant gap between the two prices indicates aggressive promotional strategies, while minimal differences suggest stable pricing policies. Additionally, incorporating the month variable allows for insights into seasonal peaks or dips in strawberry prices, such as higher prices during off-seasons or lower prices during harvest months. Together, these variables offer a comprehensive understanding of strawberry pricing dynamics in Canada’s retail market, helping inform consumer choices and vendor pricing strategies.

```{r}
#| label: fig-currentpricebyvendor
#| fig-cap: "Average Current Price by Vendor"
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_parquet(here("data","02-analysis_data","analysis_data.parquet"))
ggplot(analysis_data, aes(x = vendor, y = current_price, fill = vendor)) +
  stat_summary(fun = "mean", geom = "bar", color = "black") +
  labs(
    x = "Vendor",
    y = "Average Current Price (CAD)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The graph(@fig-currentpricebyvendor) illustrates the average current price of strawberries across various vendors, highlighting significant price variations. `Metro` stands out with the highest average price, indicating a focus on premium offerings or higher-quality products, while `No Frills` and `Save-On-Foods` have the lowest prices, appealing to cost-conscious shoppers. Vendors like `Loblaws`, `T&T`, and `Voila` fall into the mid-range pricing category, suggesting a balance between affordability and quality. These differences reflect distinct market strategies, with premium vendors targeting quality-focused consumers and budget-friendly vendors catering to price-sensitive customers. The data provides valuable insights for consumers to make informed purchasing decisions and for vendors to refine their pricing strategies to better compete in their respective market segments.

```{r}
#| label: fig-monthlytrend
#| fig-cap: "Monthly Trend of Current Prices by Vendor"
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data, aes(x = month, y = current_price, group = vendor, color = vendor)) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_point(stat = "summary", fun = "mean", size = 2) +
  labs(
    x = "Month",
    y = "Average Current Price (CAD)"
  ) +
  theme_minimal()
```

This graph(@fig-monthlytrend) shows the monthly trend of average current strawberry prices across different vendors over several months. `Metro` consistently exhibits the highest prices, with noticeable fluctuations, peaking in mid-year, indicating potential seasonal effects or changes in supply and demand. In contrast, budget-friendly vendors like `No Frills` and `Save-On-Foods` maintain the lowest prices, demonstrating pricing stability aimed at cost-conscious consumers. Mid-range vendors such as `Loblaws`, `T&T`, and `Voila` show relatively stable but slightly variable trends, with occasional peaks suggesting targeted pricing adjustments. Overall, the graph highlights clear segmentation among vendors, with premium vendors like `Metro` adapting dynamically to market conditions, while budget vendors maintain consistent affordability. This trend could reflect differing business strategies to cater to diverse consumer preferences.

## Predictor variables

@@

1. Vendor

The vendor variable is categorical, identifying the supplier of strawberries. Each vendor employs unique pricing strategies, leading to significant variability in pricing across the eight vendors (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods). Premium vendors like Metro and Galleria tend to target quality-conscious consumers, reflected in their higher average prices. In contrast, budget-oriented vendors like No Frills and Save-On-Foods maintain lower prices to attract cost-sensitive shoppers. The variability in pricing, shown through boxplots, highlights how vendor identity influences price distributions and captures the strategic segmentation of the Canadian strawberry market.

2. Old and Current Prices

@@

3. Month (month)

@@

The interplay of vendor, old price, current price, and month allows for a more comprehensive analysis of pricing strategies. For example, vendors like Metro may use larger discounts (old price vs. current price) during peak seasons to maintain competitiveness, while budget vendors like No Frills may maintain consistent pricing throughout the year. A combined analysis using bar charts and line plots highlights how vendor strategies vary with time, providing insights into the dynamic nature of strawberry pricing.

@@

# Model{#sec-model}

## Purpose of the Model

@@

- Quantify the impact of each predictor (e.g., vendor, original price, and month) on strawberry prices.
- Analyze how the gap between original (`old_price`) and discounted (`current_price`) prices varies across vendors.
- Examine seasonal trends in strawberry prices to identify periods of high or low prices.
- Provide actionable insights into vendor pricing strategies and market dynamics.

## Model Description

The chosen model is a **Bayesian linear regression model**, which builds on the simplicity of traditional linear regression while incorporating the Bayesian framework. This allows for probabilistic insights and robust uncertainty quantification, making it well-suited for analyzing the factors driving strawberry prices. Further background details and diagnostics are included in [Appendix -@sec-model-details].

**Predictors in the Model**

- **Vendor (`vendor`)**: A categorical variable identifying the supplier of strawberries. Different vendors adopt distinct pricing strategies that influence strawberry prices.

- **Original Price (`old_price`)**: A continuous variable representing the original price of strawberries before discounts or promotions. This variable reflects vendor-specific baseline pricing strategies.

- **Month (`month`)**: A numeric variable capturing the observation month, enabling the analysis of seasonal trends and time-based fluctuations.

- **Price Gap (`old_price - current_price`)**: A derived variable quantifying the magnitude of discounts or markdowns applied by vendors.

**Model set-up**

@@

\[
y_i = \beta_0 + \beta_{\text{vendor}[i]} + \beta_{\text{old\_price}} \cdot x_{\text{old\_price}[i]} + \beta_{\text{month}} \cdot x_{\text{month}[i]} + \epsilon_i
\]

@@

- \( y_i \): Current price (`current_price`) of strawberries for observation \( i \) (dependent variable).
- \( \beta_0 \): Intercept, representing the baseline price (reference vendor, average original price, and earliest month).
- \( \beta_{\text{vendor}[i]} \): Vendor-specific effect, capturing how each supplier’s prices deviate from the baseline vendor.
- \( \beta_{\text{old\_price}} \): Effect of the original price, representing how the initial pricing influences the current price.
- \( \beta_{\text{month}} \): Effect of the observation month, capturing seasonal or temporal trends in strawberry pricing.
- \( \epsilon_i \): Residual error, accounting for variability not explained by the predictors.

**Priors in the Bayesian Framework**

Coefficients (\( \beta \)):

\[
\beta \sim \text{Normal}(0, 2.5)
\]

Residual Standard Deviation (\( \sigma \)):

\[
\sigma \sim \text{Exponential}(1)
\]

The priors in this analysis reflect reasonable assumptions about the relationships between predictors and strawberry prices. Coefficients $(\beta)$ are assigned normal priors $(\operatorname{Normal}(0,2.5)$ ) to reflect the expectation that most effects are small and centered around zero, with moderate variability. For example, $\beta_{\text {old_price }}$ is expected to be positive, reflecting the influence of original prices on current prices, while $\beta_{\text {month }}$ accounts for typically small seasonal pricing effects. The intercept $\left(\beta_0\right)$ also follows a Normal $(0,2.5)$, assuming the baseline price is near zero with moderate deviations. Finally, the residual standard deviation $(\sigma)$ is given an exponential prior $(\operatorname{Exponential}(1))$ to ensure positivity and account for variability not captured by the predictors.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Model justification

A Bayesian linear regression model is a suitable approach for this analysis because it balances simplicity, interpretability, and flexibility while quantifying uncertainty in parameter estimates. This model is particularly effective for exploring the continuous outcome variable, current_price, and its relationship with predictors such as vendor, month, and original price (old price). Unlike traditional linear regression, which assumes a deterministic relationship between predictors and the outcome, the Bayesian framework provides posterior distributions for all parameters, enabling richer and more nuanced inferences. This approach allows us to incorporate prior knowledge, account for uncertainty, and provide probabilistic insights into strawberry pricing.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{rotating}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Analyzing Strawberry Pricing Strategies Across Canadian Vendors},
  pdfauthor={Tianning He},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Analyzing Strawberry Pricing Strategies Across Canadian
Vendors\thanks{Code and data are available at:
https://github.com/HeTianning/The-price-of-strawberries}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Price Variations Driven by Supplier Strategies and Seasonal
Demand in Canadian Grocery Markets}
\author{Tianning He}
\date{November 25, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study investigates the pricing of strawberries across eight major
suppliers in Canada: Voila, T\&T, Loblaws, No Frills, Metro, Galleria,
Walmart, and Save-On-Foods. Using a dataset containing variables such as
current price, old price, and month, a Bayesian regression model was
applied to analyze the relationships between pricing strategies,
seasonal trends, and vendor-specific practices. Results reveal
significant price variation among suppliers, with old price showing a
strong positive correlation with current price, reflecting the impact of
prior pricing strategies on current prices. Seasonal trends captured by
the month variable indicate subtle but consistent fluctuations in
pricing, highlighting the influence of supply and demand dynamics. These
findings provide actionable insights into supplier-specific pricing
strategies and market behavior, offering valuable guidance to consumers
and stakeholders in the Canadian strawberry market.
\end{abstract}

\section{Introduction}\label{introduction}

Strawberries are one of the most important small berries in Canada and
the United States(Macoun and McCubbin (1919)), and pricing of
strawberries in the Canadian retail market varies widely due to factors
such as supplier strategies, promotional practices, and temporal trends.
This investigation examines the determinants of strawberry prices,
focusing on differences between different suppliers, the role of
discounts, and seasonal effects. By utilizing a Bayesian linear
regression model, the study examines the relationship between key
predictors (supplier, original price, promotion, and time) and the
current price of strawberries. The Bayesian approach not only provides
parameter estimates but also quantifies uncertainty, thus providing a
nuanced understanding of price dynamics in competitive and seasonal
markets.

The main objective of this analysis is to estimate and explain the
factors that influence retail strawberry prices in Canada. The dependent
variable current price reflects the Canadian dollar price per pound of
strawberries. The key predictors include supplier identity
(categorical), which captures supplier-specific pricing strategies; old
price (continuous), which represents the original price before
promotion, and month (continuous), which captures temporal or seasonal
trends(Hodgdon et al. (2024)). By modeling these relationships, the
study aims to quantify how these variables affect pricing and explore
their combined impact on strawberry price changes.

@@

The rest of the paper is structured as follows. Section~\ref{sec-data}
describes the data collection process, including data sources, sampling
strategy, and preprocessing steps taken to ensure data quality.
Section~\ref{sec-model} outlines the methodological framework, focusing
on the Bayesian linear regression model, its formulation, and diagnostic
checks. Section~\ref{sec-result} presents the main findings of the
analysis, focusing on specific supplier pricing strategies, seasonal
trends, and the impact of sales and promotions on strawberry prices.
Section~\ref{sec-discussion} discusses the implications of these
findings, placing them in the broader context of consumer behavior and
supplier strategies in the fresh produce market.

\section{Data}\label{sec-data}

To conduct the analysis, \texttt{R}, a versatile statistical programming
language, was utilized for simulating, testing, downloading, cleaning,
and modeling data (R Core Team 2023). Several powerful libraries
supported various stages of the analysis. The \texttt{tidyverse} package
suite (Wickham et al. 2019), which includes \texttt{dplyr} for data
manipulation and \texttt{ggplot2} for visualization, was central to
handling and exploring the dataset. Data cleaning and structuring were
enhanced using \texttt{janitor}, while \texttt{arrow} facilitated
efficient handling of large datasets by enabling seamless reading and
writing of data (Apache Arrow 2021). The \texttt{here} package
streamlined file management, ensuring robust and reproducible workflows
(Müller 2020).

Visualization and reporting were further supported by \texttt{knitr}
(Xie 2021), which enabled dynamic report generation, and
\texttt{modelsummary} (Arel-Bundock 2022), which provided
professional-quality summaries for statistical models. Bayesian modeling
and advanced statistical analysis were conducted using \texttt{rstanarm}
(Goodrich et al. 2022), while posterior diagnostics and visualizations
were enhanced with \texttt{bayesplot} (Gelman, Gabry, et al. 2021).
Testing and quality assurance of code were carried out with
\texttt{testthat} (Wickham 2011), ensuring the reliability and
reproducibility of results.

@@

\subsection{Overview}\label{overview}

The data for this analysis was provided by Project Hammer (Filipp 2024),
which was collected through a process of scraping website UI screens
from vendors within the Toronto community. It contains data on product
details (e.g., product name, brand, current and old price, and unit of
product) for 8 different grocery stores (Voila, T\&T, Loblaws, No
Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods). Project
Hammer's data is available from February 28, 2024, and since the data is
collected by extracting information from grocery store websites, there
may be missing data, as discussed in appendix. Through the method of
screen scraping, the raw data is intended to measure and reflect grocery
store vendor decisions and product pricing trends.

For this analysis, the eight vendors - Voila, T\&T, Loblaws, No Frills,
Metro, Galleria, Walmart Canada, and Save-On-Foods - and their pricing
for strawberries were considered. The variable ``price'' for each
supplier in this analysis reflects the current price of their product at
the time the data was downloaded.

\subsection{Measurement}\label{measurement}

@@

The dataset also incorporates vendor as a categorical variable,
identifying one of eight major suppliers: Voila, T\&T, Loblaws, No
Frills, Metro, Galleria, Walmart, and Save-On-Foods. This variable is
pivotal for inter-vendor comparisons, enabling the analysis to uncover
unique pricing strategies, competitive behaviors, and market
segmentation. By examining differences in pricing across these
suppliers, the analysis can identify which vendors consistently offer
lower prices, adopt aggressive discounting practices, or align their
pricing with premium branding.

@@

\subsection{Outcome variable}\label{outcome-variable}

@@

By examining both current price and old price, the study provides a
detailed view of promotional activity and vendor behavior. A significant
gap between the two prices indicates aggressive promotional strategies,
while minimal differences suggest stable pricing policies. Additionally,
incorporating the month variable allows for insights into seasonal peaks
or dips in strawberry prices, such as higher prices during off-seasons
or lower prices during harvest months. Together, these variables offer a
comprehensive understanding of strawberry pricing dynamics in Canada's
retail market, helping inform consumer choices and vendor pricing
strategies.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-currentpricebyvendor-1.pdf}

}

\caption{\label{fig-currentpricebyvendor}Average Current Price by
Vendor}

\end{figure}%

The graph(Figure~\ref{fig-currentpricebyvendor}) illustrates the average
current price of strawberries across various vendors, highlighting
significant price variations. \texttt{Metro} stands out with the highest
average price, indicating a focus on premium offerings or higher-quality
products, while \texttt{No\ Frills} and \texttt{Save-On-Foods} have the
lowest prices, appealing to cost-conscious shoppers. Vendors like
\texttt{Loblaws}, \texttt{T\&T}, and \texttt{Voila} fall into the
mid-range pricing category, suggesting a balance between affordability
and quality. These differences reflect distinct market strategies, with
premium vendors targeting quality-focused consumers and budget-friendly
vendors catering to price-sensitive customers. The data provides
valuable insights for consumers to make informed purchasing decisions
and for vendors to refine their pricing strategies to better compete in
their respective market segments.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-monthlytrend-1.pdf}

}

\caption{\label{fig-monthlytrend}Monthly Trend of Current Prices by
Vendor}

\end{figure}%

This graph(Figure~\ref{fig-monthlytrend}) shows the monthly trend of
average current strawberry prices across different vendors over several
months. \texttt{Metro} consistently exhibits the highest prices, with
noticeable fluctuations, peaking in mid-year, indicating potential
seasonal effects or changes in supply and demand. In contrast,
budget-friendly vendors like \texttt{No\ Frills} and
\texttt{Save-On-Foods} maintain the lowest prices, demonstrating pricing
stability aimed at cost-conscious consumers. Mid-range vendors such as
\texttt{Loblaws}, \texttt{T\&T}, and \texttt{Voila} show relatively
stable but slightly variable trends, with occasional peaks suggesting
targeted pricing adjustments. Overall, the graph highlights clear
segmentation among vendors, with premium vendors like \texttt{Metro}
adapting dynamically to market conditions, while budget vendors maintain
consistent affordability. This trend could reflect differing business
strategies to cater to diverse consumer preferences.

\subsection{Predictor variables}\label{predictor-variables}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Vendor
\end{enumerate}

The vendor variable is categorical, identifying the supplier of
strawberries. Each vendor employs unique pricing strategies, leading to
significant variability in pricing across the eight vendors (Voila,
T\&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods).
Premium vendors like Metro and Galleria tend to target quality-conscious
consumers, reflected in their higher average prices. In contrast,
budget-oriented vendors like No Frills and Save-On-Foods maintain lower
prices to attract cost-sensitive shoppers. The variability in pricing,
shown through boxplots, highlights how vendor identity influences price
distributions and captures the strategic segmentation of the Canadian
strawberry market.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Old and Current Prices
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Month (month)
\end{enumerate}

@@

The interplay of vendor, old price, current price, and month allows for
a more comprehensive analysis of pricing strategies. For example,
vendors like Metro may use larger discounts (old price vs.~current
price) during peak seasons to maintain competitiveness, while budget
vendors like No Frills may maintain consistent pricing throughout the
year. A combined analysis using bar charts and line plots highlights how
vendor strategies vary with time, providing insights into the dynamic
nature of strawberry pricing.

@@

\section{Model}\label{sec-model}

\subsection{Purpose of the Model}\label{purpose-of-the-model}

@@

\begin{itemize}
\tightlist
\item
  Quantify the impact of each predictor (e.g., vendor, original price,
  and month) on strawberry prices.
\item
  Analyze how the gap between original (\texttt{old\_price}) and
  discounted (\texttt{current\_price}) prices varies across vendors.
\item
  Examine seasonal trends in strawberry prices to identify periods of
  high or low prices.
\item
  Provide actionable insights into vendor pricing strategies and market
  dynamics.
\end{itemize}

\subsection{Model Description}\label{model-description}

The chosen model is a \textbf{Bayesian linear regression model}, which
builds on the simplicity of traditional linear regression while
incorporating the Bayesian framework. This allows for probabilistic
insights and robust uncertainty quantification, making it well-suited
for analyzing the factors driving strawberry prices. Further background
details and diagnostics are included in
Appendix~\ref{sec-model-details}.

\textbf{Predictors in the Model}

\begin{itemize}
\item
  \textbf{Vendor (\texttt{vendor})}: A categorical variable identifying
  the supplier of strawberries. Different vendors adopt distinct pricing
  strategies that influence strawberry prices.
\item
  \textbf{Original Price (\texttt{old\_price})}: A continuous variable
  representing the original price of strawberries before discounts or
  promotions. This variable reflects vendor-specific baseline pricing
  strategies.
\item
  \textbf{Month (\texttt{month})}: A numeric variable capturing the
  observation month, enabling the analysis of seasonal trends and
  time-based fluctuations.
\item
  \textbf{Price Gap (\texttt{old\_price\ -\ current\_price})}: A derived
  variable quantifying the magnitude of discounts or markdowns applied
  by vendors.
\end{itemize}

\textbf{Model set-up}

@@

{[} y\_i = \beta\emph{0 + \beta}\{\text{vendor}{[}i{]}\} +
\beta\emph{\{\text{old\_price}\} \cdot x}\{\text{old\_price}{[}i{]}\} +
\beta\emph{\{\text{month}\} \cdot x}\{\text{month}{[}i{]}\} +
\epsilon\_i {]}

@@

\begin{itemize}
\tightlist
\item
  \(y_i\): Current price (\texttt{current\_price}) of strawberries for
  observation \(i\) (dependent variable).\\
\item
  \(\beta_0\): Intercept, representing the baseline price (reference
  vendor, average original price, and earliest month).\\
\item
  \(\beta_{\text{vendor}[i]}\): Vendor-specific effect, capturing how
  each supplier's prices deviate from the baseline vendor.\\
\item
  \(\beta_{\text{old\_price}}\): Effect of the original price,
  representing how the initial pricing influences the current price.\\
\item
  \(\beta_{\text{month}}\): Effect of the observation month, capturing
  seasonal or temporal trends in strawberry pricing.\\
\item
  \(\epsilon_i\): Residual error, accounting for variability not
  explained by the predictors.
\end{itemize}

\textbf{Priors in the Bayesian Framework}

For the coefficients (\(\beta\)):

{[} \beta \sim \text{Normal}(0, 2.5) {]}

For the residual standard deviation (\(\sigma\)):

{[} \sigma \sim \text{Exponential}(1) {]}

The priors in this analysis reflect reasonable assumptions about the
relationships between predictors and strawberry prices. Coefficients
\((\beta)\) are assigned normal priors \((\operatorname{Normal}(0,2.5)\)
) to reflect the expectation that most effects are small and centered
around zero, with moderate variability. For example,
\(\beta_{\text {old_price }}\) is expected to be positive, reflecting
the influence of original prices on current prices, while
\(\beta_{\text {month }}\) accounts for typically small seasonal pricing
effects. The intercept \(\left(\beta_0\right)\) also follows a Normal
\((0,2.5)\), assuming the baseline price is near zero with moderate
deviations. Finally, the residual standard deviation \((\sigma)\) is
given an exponential prior \((\operatorname{Exponential}(1))\) to ensure
positivity and account for variability not captured by the predictors.

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of Goodrich et al. (2022). We use the default priors from
\texttt{rstanarm}.

\subsection{Model justification}\label{model-justification}
